Detailed structure of phospholipid bilayers with negatively charged lipids and salt concentrations from molecular dynamics simulations including the effects of electronic polarization

### 2019-05-17 Manuscript/manuscript.tex
Edits in Results section one (form fact. etc.)

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{NMRlipidsIV}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
\end{figure}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/dihedral_angle_distribution_Ca-Cb-Cg-Og_ECC-L17.pdf}
  \caption{\label{fig:dihedral}
    Distribution of dihedral angle \ce{C_\alpha-C_\beta-C_\gamma-O_\gamma} in POPS 
    from simulations of a pure POPS bilayer with \ce{Na+} counterions using Lipid17 and ECC-lipids. 
}
\end{figure}

@@

Recent analyzes of interaction sites between calcium and PS lipids combining
MD simulations with various experimental techniques have been 
controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
\todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
However, in CHARMM36 simulations without the NBfix correction for calcium,
calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
was not observed in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
PS lipid fraction and calcium concentration was used in previous work \cite{hallock18},
which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
%While POPC interacts with the calcium cations almost entirely through its phosphate group 
%in both neutral \cite{melcr18} and negatively charged membranes, 
%the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
%Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
%however, they are always accompanied by interactions with phosphate groups. 
Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
%This is in contrast to 
%combined with fluorescent and vibrational sum frequency spectroscopy,
%which suggest a 
Overestimated coordination of cations with carbonyl groups is a potential source for
overestimated cation binding in Berger force fields \cite{catte16,NMRlipidsIV}. 
In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.

@@

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidthsmall]{../img/bound_ca_populations.pdf} 
  \includegraphics[width=\figwidthsmall]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
  \caption{ \label{fig:cacl_complexes} 
  Left: Percentages of the population of bound Ca$^{2+}$ 
in exclusive contact with various lipid moieties 
in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} \\
    Right: Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
    Analysis was done for both systems by considering all lipids (blue and violet) and
    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). \\
    %All lipids were taken into account with the exception of the complexes in light green and orange, 
    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
    Clusters of four or more lipids were not observed in either membrane.
    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
    as the distance between the cation and the oxygen atoms of the lipids. 
    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
  }
\end{figure}

>>>>

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from the same simulations 
    compared with experiments at 298~K. \citep{NMRlipidsIV}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
\todo{Add label $\gamma$ to the figure -- it is used in Fig with the dihedral angle, 
and it can be also used in the description of COO- vector. }
\end{figure}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/dihedral_angle_distribution_Ca-Cb-Cg-Og_ECC-L17.pdf}
  \caption{\label{fig:dihedral}
    Distribution of dihedral angle \ce{C_\alpha-C_\beta-C_\gamma-O_\gamma} in POPS 
    from simulations of a pure POPS bilayer with \ce{Na+} counterions using Lipid17 and ECC-lipids. 
    Labelling of POPS segments is in Fig.~\ref{simVSexpNOions_POPS}. 
}
\end{figure}

@@

Recent analyzes of interaction sites between calcium and PS lipids combining
MD simulations with various experimental techniques have been 
controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
the calcium ions bind more likely 
to the carboxylate than to phosphate moiety of POPS headgroup (Table \ref{tab:binding}),
in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
was not observed in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
PS lipid fraction and calcium concentration was used in previous work \cite{hallock18},
which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
%While POPC interacts with the calcium cations almost entirely through its phosphate group 
%in both neutral \cite{melcr18} and negatively charged membranes, 
%the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
%Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
%however, they are always accompanied by interactions with phosphate groups. 
Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
%This is in contrast to 
%combined with fluorescent and vibrational sum frequency spectroscopy,
%which suggest a 
Overestimated coordination of cations with carbonyl groups is a potential source for
overestimated cation binding in Berger force fields \cite{catte16,NMRlipidsIV}. 
In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.

@@

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidthsmall]{../img/bound_ca_populations.pdf} 
  \includegraphics[width=\figwidthsmall]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
  \caption{ \label{fig:cacl_complexes} 
  \textbf{Left:} Percentages of the population of bound Ca$^{2+}$ 
in exclusive contact with various lipid moieties 
in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
The bar representing total populations of bound \ce{Ca^{2+}} to PC, PS, or PC and PS (blue),
is in part covered by areas representing contributions of individual moieties (shades of orange and red). 
In the case of mixed simultaneous binding to PC and PS
we distinguish different moieties only in PS.
%i.e., PC can interact with any considered moiety. 
For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
    \textbf{Right:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
    Analysis was done for both systems by considering all lipids (blue and violet) and
    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
    %All lipids were taken into account with the exception of the complexes in light green and orange, 
    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
    Clusters of four or more lipids were not observed in either membrane.
    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
    as the distance between the cation and the oxygen atoms of the lipids. 
    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
  }
\end{figure}
